NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2019-02-05 Manuscript/manuscriptPS.tex

However, the best performing models,
Slipids, CHARMM36 and CHARMM36ua, reproduce the larger forking of the $\alpha$-carbon 
and the Slipids model reproduces also the lower of the $\beta$-carbon order parameter
when comparing the PS results to PC (Fig. \ref{HGorderParametersPS} vs. Fig. 2 in Ref.~\citenum{botan15}).
Interestingly, the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$  dihedral with
a single and narrow peak in the angle distribution close to 120$^{\circ}$ is more restricted
in the best three models than in other models which give two maxima with different angles (Fig. \ref{dihedralsHG}).
The restricted motion is also
visible in the sampled conformations (Figs.~\ref{HGstructuresPSandPC} (A) and \ref{HGandGLYstructuresPS})
suggesting that the rotation of carboxyl group is limited in the serine headgroup.
In addition, the N-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral exhibit a more asymmetric
and restricted angle distribution for PS than for PC headgroup in
CHARMM36 simulations in best agreement with experiments
(Figs.~\ref{HGstructuresPSandPC} (B) and \ref{dihedralsHGpc}).
The results might manifest the increased rigidity anticipated in the early experimental studies \cite{browning80,buldt81}.
Also, the sampled conformations of glycerol backbone
significantly vary between different simulation models 
(Figs. \ref{dihedralsGLY}, \ref{HGandGLYstructuresPS}, \ref{dihedralsGLYpc} and \ref{HGandGLYstructuresPSPC}),
but further analysis is beyond the scope of this work which the PS headgroup.

The suggested characteristic conformations of PS headgroup may be useful when interpreting experiments and
can guide the further force field development. However, more accurate simulations are
required to confirm the suggested conformations because the simulated PS headgroup order parameters
are not within the experimental error bars in any of the tested models.

>>>>

However, if one excludes the glycerol backbone and focuses only on the headgroup ($\alpha$ and $\beta$) order parameters,
the best performing models,
Slipids, CHARMM36 and CHARMM36ua, do reproduce the larger-than-in-PC forking of the $\alpha$-carbon.
Slipids reproduces also the feature that the $\beta$-carbon order parameter
in PS is significantly smaller than in PC (Fig. \ref{HGorderParametersPS} vs. Fig. 2 in Ref.~\citenum{botan15}).
Interestingly, the C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$  dihedral with
a single and narrow peak in the angle distribution close to 120$^{\circ}$ is more restricted
in the best three models than in other models which give two maxima with different angles (Fig. \ref{dihedralsHG}).
The restricted motion is also
visible in the sampled conformations (Figs.~\ref{HGstructuresPSandPC} (A) and \ref{HGandGLYstructuresPS})
suggesting that the rotation of the carboxyl group is limited in the serine headgroup.
In addition, the N-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral exhibits a more asymmetric
and restricted angle distribution for PS than for PC headgroup in
CHARMM36 simulations in best agreement with experiments
(Figs.~\ref{HGstructuresPSandPC} (B) and \ref{dihedralsHGpc}).
The results might manifest the increased rigidity anticipated in the early experimental studies \cite{browning80,buldt81}.
Also, the sampled conformations of glycerol backbone
significantly vary between different simulation models 
(Figs. \ref{dihedralsGLY}, \ref{HGandGLYstructuresPS}, \ref{dihedralsGLYpc} and \ref{HGandGLYstructuresPSPC}),
but further analysis is beyond the scope of this work which focuses on the PS headgroup.

The suggested characteristic conformations of the PS headgroup may be useful when interpreting experiments and
can guide further force field development. However, more accurate MD force fields are
required to confirm the suggested conformations because the simulated PS headgroup order parameters
are not within the experimental error in any of the tested models.
